### 2014-02-28 bibliography.tex

\begin{thebibliography}{9}
\bibitem{codd}
  \textit{Codd E.F., Codd S.B., Salley C.T.}
  Providing OLAP (On-line Analytical Processing) to User-Analysts: An IT
  Mandate. – Sunnyvale(CA): Codd \& Date Inc.,
  1993.
  – 31 p.
\bibitem{lecht}
  \textit{Lechtenborger J., Vossen G.}
  Multidimensional normal forms for data warehouse design
  // Inf. Syst.
  – 2003.
  – Vol. 28, N 5. – P. 415–434. 
\bibitem{lehner}
  \textit{Lehner W., Albrecht J., Wedekind H.}
  Normal forms for multidimensional databases
  // Proc. of the Tenth Intern. Conf. on Scientific and Statistical Database
  Management.
  – Capri, 1998.
  – P. 63–72.
\bibitem{mazon}
  \textit{Mazon J.-N., Trujillo J., Lechtenborger J.}
  Reconciling requirement-driven data warehouses with data sources via
  multidimensional normal forms
  // Data Knowledge Engineering.
  – 2007.
  – Vol. 63, N 3. – P. 725–751.
\bibitem{vassi}
  \textit{Vassiliadis P., Sellis T.}
  A survey of logical models for OLAP databases
  // SIGMOD Rec.
  – 1999.
  – Vol. 28, N 4. – P. 64–69.
\bibitem{peder}
  \textit{Pedersen T.B., Jensen C.S., Dyreson C.E.}
  A foundation for capturing and querying complex multidimensional data
  // Inf. Syst.
  – 2001.
  – Vol. 26, N 5. – P. 383–423.
\bibitem{progressive}
  \textit{Progressive ranking of range aggregates}
  / H.-G. Li, et al.
  // Data \& Knowledge Engineering.
  – 2007.
  – Vol. 63, N 1. – P. 4–25. 
\bibitem{giorg}
  \textit{Giorgini P., Rizzi S., Garzetti M.}
  Goal-oriented requirement analysis for data warehouse design
  // Proc. of the 8th ACM international Workshop on Data Warehousing and OLAP: DOLAP '05.
  – Bremen, 2005.
  – P. 47–56.
%ours  
\bibitem{zyk_pol}
  \textit{С.В. Зыкин, А.Н. Полуянов}
  Формирование многомерных данных с использованием промежуточных представлений.
  "Проблемы управления".
  2013.
  № 5. Стр. 54-59.
%stream
\bibitem{Olston03}
  \textit{Olston C., Jiang J., Widom J.}
  Adaptive filters for continuous queries over distributed data streams
  // Proc. of the 2003 ACM SIGMOD Intern. Conf. on Management of Data (SIGMOD '03).
  – San Diego, 2003.
  – P. 563–574.
\bibitem{Denny05}
  \textit{Denny M., Franklin M.J.}
  Predicate result range caching for continuous queries
  // Proc. of the 2005 ACM SIGMOD Intern. Conf. on Management of Data (SIGMOD '05).
  – N.-Y., 2005.
  – P. 646–657.
%logic
\bibitem{Afrati06}
  \textit{Afrati F., Li C., Mitra P.}
  Rewriting queries using views in the presence of arithmetic comparisons
  // Theoretical Computer Science.
  – 2006.
 – Vol. 368, N 1–2. – P. 88–123.
%static approach
\bibitem{baralis}
  \textit{Baralis, E., Paraboschi, S., Teniente, E.} Materialized view selection in a multidimensional database.
  // Proc. of the 23rd International Conference on Very Large Data Bases,
  Athens, Greece.
  - 1997
  - P. 318–329
\bibitem{gupta}
  \textit{Gupta, H.}
  Selection of views to materialize in a data warehouse.
  // Proceedings of the International Conference on Database Theory, Delphi.
  - Greece, 1997.
  - P. 98–112.
\bibitem{gupta-mumick}
  \textit{Gupta, H., Mumick, I.S.}
  Selection of views to materialize under a maintenance cost constraint.
  // Proceedings of the International Conference on Database Theory.
  - Israel, 1999.
  - P. 453–470.
%dynamic
\bibitem{scheuermann}
  \textit{Scheuermann, P., Shim, J., Vingralek R.}
  WATCHMAN: A data warehouse intelligent cache manager.
  // Proceedings of the 22nd International Conference on Very Large Data Bases.
  - Bombay, India, 1996.
  - P. 51–62.
\bibitem{shim}
  \textit{Shim, J., Scheuermann, P., Vingralek R.}
  Dynamic caching of query results for decision support systems.
  // Proceedings of the 11th International Conference on Scientific and
  Statistical Database Management.
  - Cleveland, OH, 1999.
  - P. 254–263.
\bibitem{kalnis}
  \textit{Kalnis, P., Papadias, D.}
  Proxy-server architectures for OLAP.
  // Proceedings of the 2001 ACM SIGMOD International Conference on Management
  of Data.
  - Santa Barbara, CA, 2001.
  - P. 367–378.
\bibitem{chang}
  \textit{Chang-Sup Park, Myoung Ho Kim b, Yoon-Joon Lee}
  Usability-based caching of query results in OLAP systems.
  // The Journal of Systems and Software
  - 2003
  - Vol. 68, - P. 103–119
 %domain
\bibitem{Keller96}
  \textit{Keller M., Basu J.}
  A predicate-based caching scheme for client-server database architectures
  // VLDB Journal. – 1996. N 5. – P. 35–47.

\end{thebibliography}